---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache=TRUE
)


```

# EC 

## Overview

This package calculates the exact-corrected confidence interval [Hawila & Berg, 2021][HB21] and other confidence intervals for noninferiorty binomial trials with various methods. Additionally, as described in [Hawila & Berg (2021)][HB21], functions are provided to calculate maximal sizes and p-values corresponding to different confidence interval methods.


## Installation

```{r, eval = FALSE}
library(devtools)
install_github("NourHawila/EC")
```

## Load the package

```{r}
library(EC)
```

## Usage

We consider a noninferiority trial with treatment group ($T$) and control/standard-of-care group ($C$) having a binary endpoint representing whether or not an outcome is observed. Here we assume the outcome is a positive event  (e.g. resolution of a disease).

Let $P_T$ and $P_C$ be the probabilities the outcome is observed, and let $\delta=P_T-P_C$ represent the risk difference. We consider the following hypotheses:

$H_0: \delta\le -\delta_0$: "inferior trial"; $T$ is inferior to $C$ 

$H_1: \delta> -\delta_0$: "non-inferior trial"; $T$ is not inferior to $C$

Let `x.T`, `x.C`, `N.T`, `N.C`, `delta0`, and `alpha` be the user-defined values for the noninferiority trial. 

We use small sample sizes in the example below as the Chan & Zhang confidence interval method can take a while to compute.

```{r}
x.T=5
x.C=2
N.T=6
N.C=6  
delta0=.12
alpha=.05
```

### Confidence intervals

Note that only the exact corrected confidence interval, `ci_EC`, uses the noninferiority margin $\delta_0$ in constructing the confidence interval. Note that $H_0$ is rejected if the lower bound of the confidence interval is larger than $-\delta_0$.

```{r}
ci_EC(x.T=x.T,x.C=x.C,N.T=N.T,N.C=N.C,alpha=alpha,delta0=delta0)
ci_CZ(x.T=x.T,x.C=x.C,N.T=N.T,N.C=N.C,alpha=alpha)
ci_MN(x.T=x.T,x.C=x.C,N.T=N.T,N.C=N.C,alpha=alpha)
ci_Wald(x.T=x.T,x.C=x.C,N.T=N.T,N.C=N.C,alpha=alpha)
```


### p-values

As described in [Hawila & Berg, 2021][HB21], we correspond a p-value with each confidence interval method. Note that $H_0$ is rejected if the p-value is smaller than $\alpha/2$.

```{r}

pval_Chan(x.T=x.T,x.C=x.C,N.T=N.T,N.C=N.C,delta0=delta0)
pval_CZ(x.T=x.T,x.C=x.C,N.T=N.T,N.C=N.C,delta0=delta0)
pval_MN(x.T=x.T,x.C=x.C,N.T=N.T,N.C=N.C,delta0=delta0)
pval_Wald(x.T=x.T,x.C=x.C,N.T=N.T,N.C=N.C,delta0=delta0)

```

### Maximal sizes

As described in [Hawila & Berg, 2021][HB21], for a given value of $N_T$, $N_C$, and $\alpha$, we can also calculate the maximal size for each confidence interval method. The size of the exact-corrected method is bounded by $\alpha/2$, but the sizes of the other methods can be larger than $\alpha/2$.

```{r}
size_EC(alpha=alpha,N.T=N.T,N.C=N.C,delta0=delta0)
#size_CZ(alpha=alpha,N.T=N.T,N.C=N.C,delta0=delta0)
size_MN(alpha=alpha,N.T=N.T,N.C=N.C,delta0=delta0)
size_Wald(alpha=alpha,N.T=N.T,N.C=N.C,delta0=delta0)
```

## Getting help

```{r, eval = FALSE}
help(package="EC")
```



[HB21]: https://arxiv.org/pdf/2104.04660.pdf